---
title: "Chapter 1. Review of Linear Regression"
author: 
- JoonHo Lee (jlee296@ua.edu)
- Mehdi Rajeb
date: "October 5, 2021"
output:
  html_document: 
    css: styles.css
    fig_caption: yes
    highlight: haddock
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 1
  tufte::tufte_html:
    number_sections: yes
    toc: true
    toc_depth: 2
  pdf_document:
    toc: no
    toc_depth: '3'
    number_sections: yes
bibliography: bibliography.bib
---

```{r basic_setup, include=FALSE}
### Set working directory and data directory 
work_dir <- file.path(path.expand("~"), 
                      "Documents",
                      "RStan-companion-for-MLMUS4") 

data_dir <- file.path(work_dir, "datasets")

setwd(work_dir)

### Set R Markdown options
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, message = FALSE)

### Call libraries
library(tidyverse)
library(haven)
library(broom)
library(janitor)

### Define manual palettes
color_palette <- c("firebrick1", "dodgerblue1", "forestgreen", "darkorchid1",
                   "darkgoldenrod1", "blue", "green", "purple", "gold", "red")    
shape_palette <- c(16, 17, 15, 18, 1, 2, 0, 5, 6, 4, 3, 8, 10, 7, 9) 
```


# Introduction

In this chapter, we review the statistical models underlying independent-samples $t$ tests, analysis of variance (ANOVA), analysis of covariance (ANCOVA), simple
regression, and multiple regression. We formulate all these models as linear regression
models.


# Is there gender discrimination in faculty salaries?

The data considered here are a subset of the data provided by DeMaris, comprising $n = 514$ faculty members, excluding faculty from the Fireland campus, nonprofessors
(instructors/lecturers), those not on graduate faculty, and three professors hired
as Ohio Board of Regents Eminent Scholars. We will use the following variables:

- `salary`: academic year (9-month) salary in U.S. dollars
- `male`: gender (1 = male; 0 = female)
- `market`: marketability of academic discipline, defined as the ratio of the national average salary paid in the discipline to the national average across all disciplines
- `yearsdg`: time since degree (in years)
- `rank`: academic rank (1 = assistant professor; 2 = associate professor; 3 = full professor)

We start by reading the data into R:

```{r}
# Import faculty dataset
load_path <- file.path(data_dir, "faculty.dta")
df <- read_dta(load_path)
df

# Covert numerics into factors
df <- df %>%
  mutate(
    rank = factor(rank, 
                  levels = c(1, 2, 3), 
                  labels = c("Assistant", "Associate", "Full")), 
    male = factor(male, 
                  levels = c(0, 1), 
                  labels = c("Women", "Men"))
  )
```


# Independent-samples $t$ test

```{r}
df_sum <- df %>%
  group_by(male) %>%
  summarise(
    Mean = mean(salary, na.rm = TRUE), 
    SD = sd(salary, na.rm = TRUE), 
    N = n()
  )

df_sum

# df_total <- df %>%
#     summarise(
#     Mean = mean(salary), 
#     SD = sd(salary), 
#     N = n()
#   ) %>%
#   add_column(male = "Total", .before = "Mean")
# 
# df_sum
# df_total

```


```{r}
df %>%
  ggplot(aes(y = salary)) + 
  geom_boxplot() + 
  facet_wrap(. ~ male) +
  theme_bw() + 
  labs(y = "Academic Salary (in dollars)")
```



```{r}
t_test <- t.test(salary ~ male, df)
t_test
broom::tidy(t_test)
```






# References








